Covariate analysis of MAYO reprocessed counts (with CQN normalisation and covariates adjusted)

### 2017-09-05 code/Rmd/MAYO_REPROCESSED.Rmd
Minor modifications, added AOD specific differential expression to MAYO markdown

>>>>

### Differential expression analysis (with BrainRegion.Diagnosis x AgeAtDeath as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp3, fig.height=5, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('BrainRegion.Diagnosis', setdiff(postAdjustCovars, 'AgeAtDeath')), drop = F], 
                         Intercept = F)
DESIGN$design[,grep('BrainRegion.Diagnosis',colnames(DESIGN$design))] = DESIGN$design[,grep('BrainRegion.Diagnosis',colnames(DESIGN$design))] * COVARIATES[,'AgeAtDeath']
colnames(DESIGN$design)[grep('BrainRegion.Diagnosis',colnames(DESIGN$design))] = 
  paste(colnames(DESIGN$design)[grep('BrainRegion.Diagnosis',colnames(DESIGN$design))],'AOD', sep = '.')
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]

# Fit contrast
contrast = makeContrasts(contrasts=c(
  "BrainRegion.DiagnosisCER.AD.AOD-BrainRegion.DiagnosisCER.Control.AOD",
  "BrainRegion.DiagnosisCER.AD.AOD-BrainRegion.DiagnosisCER.OTHER.AOD",
  "BrainRegion.DiagnosisCER.OTHER.AOD-BrainRegion.DiagnosisCER.Control.AOD",
  "BrainRegion.DiagnosisTCX.AD.AOD-BrainRegion.DiagnosisTCX.Control.AOD",
  "BrainRegion.DiagnosisTCX.AD.AOD-BrainRegion.DiagnosisTCX.OTHER.AOD",
  "BrainRegion.DiagnosisTCX.OTHER.AOD-BrainRegion.DiagnosisTCX.Control.AOD"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('BrainRegion.Diagnosis','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('BrainRegion.ref','Dx.ref','AOD.ref'), sep = '\\.') %>%
  tidyr::separate(to.state, c('BrainRegion.ag','Dx.ag','AOD.ag'), sep = '\\.') %>%
  tidyr::unite(Comparison, Dx.ref, Dx.ag, sep = '-') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gc_content, gene.length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'

tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction, BrainRegion.ref, AOD.ref) %>%
  group_by(BrainRegion.ref, AOD.ref, Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)

p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Comparison+.~BrainRegion.ref+AOD.ref, scales = 'fixed')
p

all.diff.exp = c(all.diff.exp, list(Diagnosis.AOD = DE))
all.fit = c(all.fit, list(Diagnosis.AOD = FIT))
```
